---
title: "MATH1062 Lab6 Worksheet"
subtitle: "Introduction to R"
author: "© University of Sydney MATH1062"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---

**Learning Outcome** 

LO1. apply mathematical logic and statistical thinking to solve problems
    
LO5. apply statistical software such as R to analyse example sets of data

At the end of this Lab, you should be able to:

- Understand the relationship between **population** and **sample** via simulation
- Simulating probability using R
- Being able to use factorial and combinations to work out exact probabilities 
- Make the result reproducible by fixing random seed 


# Simulating Chance

In Semester 2 2022, MATH1005 had 928 students. In the MATH1005 survey data set, we only have 281 observations.
Hence, we only have a sample of the  MATH1005 cohort  in S2, 2022.

Using a simulation we will solve the following exercise:

- **Assume that, out of the 928 students, 427 are female. When picking randomly 281 students out of 928, what is the probability that more than 50% of the sample are female?**

To do this, implement the following steps:

- Use the ``sample`` function to draw without replacement a sample of 281 from a population with 427 `1`'s and 501 `0`'s.  Call it ``my.sample``.
    - *Hint 1: Use ``?sample`` to open the help file of the sample function*.
    - *Hint 2: The population (x) can be written as ``c(rep(1,427), rep(0,501))``*.
    
```{r}
### modify the following code
# here is my example of a different population with 500 1s and 500 0s
population = c(rep(1,500), rep(0,500))
# here is my example of drawing 200
my.sample = sample(population,200)
```

- Apply the ``mean`` function on ``my.sample``. This gives you the proportion of `1`'s (= females) in your draw.  Call the result ``prop.female``.

```{r}
### Write your code here
set.seed(0) # does not change
poulation = c(rep(1,427),rep(0,928-427))
my.sample = sample(population,281,replace = FALSE)
```

```{r}
prop.female = mean(my.sample) # femal is the one
```

- Write ``female.5 = prop.female > 0.5``. The variable `female.5` tells you whether more than 50% are female in the draw (`TRUE`) or not (`FALSE`).

```{r}
### Write your code here
female.5 = prop.female > 0.5
female.5
```

- Create a function for one draw:

```{r}
drawing = function(){
  ### Write your code here
  poulation = c(rep(1,427),rep(0,928-427))
  my.sample = sample(population,281,replace = FALSE)
  prop.female = mean(my.sample)
  female.5 = prop.female > 0.5
  return(female.5)
}           
```

- Type `drawing()` and get the result of one draw. You can press the "Run current chunk" button several times.

```{r}
drawing()
```
- It seems that `FALSE` (i.e. less than 50% are female) happens quite often.

- Use the ``replicate`` function to repeat the draw 10,000 times. Call the result ``drawings`` (type  `` drawings = replicate(10000,drawing()) `` )
- Note: you can use `set.seed` function to make your results reproducible. Why are the results not reproducible otherwise?

```{r}
### Write your code here
set.seed(10)
data = replicate(1000,drawing())
data
```

- Take the mean of ``drawings`` via the ``mean`` function. 
```{r}
### Write your code here
prop = mean(data)
prop
```
- The result is the desired probability. That is, **from a population of 928 students, where 427 are female, a random sample of 281 students has approximately this probability of having more than 50% females**.


# Factorial and sampling with/without replacement

The “bootstrap” is a statistical method for generating a new data set that is like an existing one. Suppose we have a data set consisting of 6
observations: 
\[
\{x_1,x_2,x_3,x_4,x_5,x_6\}. 
\]
To generate a “bootstrap” data set, we sample from the original data set with replacement, meaning that it is possible for each observation to be sampled more than once. Examples of bootstrap data sets include:
\[\{x4,x2,x4,x3,x2,x4\}\]
\[\{x3,x1,x6,x1,x1,x2\}\]
\[\{x2,x1,x4,x3,x6,x5\}\] 
Notice that in the last example, each observation appears exactly once (which corresponding to the outputs of sampling without replacement). What is the probability that the bootstrap data set contains each observation exactly once?

**Applying what we learned in lectures, we want to derive the exact probability and simulate the probability in R. **

In R simulation, the following function is useful to check if a “bootstrap” data set contains each observation exactly once.

```{r}
# you can modify this function in your simulation
check_sample <- function(bootstrap_sample){
  # returns TRUE if the bootstrap_sample contains each observation exactly once
  # returns FALSE otherwise
  #
  # check if the frequencies of all categories in the frequency table is 1
  x = all(table(bootstrap_sample) == 1)
  return(x)
}

# here are examples of calling this function
bootstrap_sample <- sample(c("x1", "x2", "x3", "x4", "x5", "x6"), size = 6, replace = TRUE)
check_sample(bootstrap_sample)
#
bootstrap_sample <- sample(c("x1", "x2", "x3", "x4", "x5", "x6"), size = 6, replace = FALSE)
check_sample(bootstrap_sample)
```

Now it is your turn to write your own code to simulate the bootstraping process many times (maybe $\geq 10^5$) and count the proportion of bootstrap data set containing each observation exactly once.

```{r}
### Write your code here

# see below the solution
set.seed(5)
sample_space = c("x1", "x2", "x3", "x4", "x5", "x6")
res = replicate(10000,check_sample(sample(sample_space,6,replace = TRUE)))

mean(res)
```


# Combinations

We toss a fair coin 20 times. Each toss results in two equally likely outcomes, head or tail. 

- What is the probability that you get exactly 10 heads in the 20 tosses?

- What is the probability that you get exactly 5 heads in the 20 tosses?

**Derive the exact probability (without directly using Binomial distribution if you know what is that) and simulate the probability in R. **

```{R}
### Write your code here
# see the end of the file
sample_space = c(1,0)

result = replicate(10000,sum(sample(sample_space,20,replace = TRUE)) == 10)

mean(result)

result = replicate(10000,sum(sample(sample_space,20,replace = TRUE)) == 5)

mean(result)
```
