---
title: "MATH1062 Lab3 Worksheet"
subtitle: "Introduction to R"
author: "© University of Sydney MATH1062"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

# <i class="fa fa-crosshairs"></i> Aim of Lab3 {-}

**Learning Outcome** 

LO1. apply mathematical logic and statistical thinking to solve problems

LO3. identify appropriate methods to describe, summarise and visualise a given data set

LO5. apply statistical software such as R to analyse example sets of data

By now, you should already be able to:

- Import a given data-set into R
- Visualise variables by choosing an appropriate graphical summary


At the end of this Lab, you should be able to:

- Present and interpret graphical/numerical summaries of variables and understand their limitations
- Write a function in R
- Model a variable by a Normal curve, and explain how to find areas under the curve and percentiles.
- Use R to calculate areas under the Normal curve and percentiles.

# Import data and IDA


- Save the data-file `math1005_cleaned.csv`  in your data folder within your MATH1062 folder.

- **Import** the data into a variable called `math1005`:

```{r}
math1005 = read.csv("math1005_cleaned.csv")
```

This data was collected from an **optional survey** for MATH1005 students at the start of 2022 Semester 2. There are a total of **928 students** enrolled in this course.

- How many survey **responses** are there? How many **variables** are there?

```{r}
### write code here
dim(math1005)
```

- What are the **columns** of your data and what are their **type**? Are they classified correctly?
    - *Hint: Use the function `str `. Other useful functions are `names` and`head`*.
     
```{r}
### write code here
str(math1005)
math1005$Country
```


# Quantitative Data

## Height

- Create a **graphical summary** for the **height** distribution of the cohort. Comment on the distribution, are there any **outliers**? Are there any issues with the data?

```{r}
### write code here

boxplot(math1005$Height,horizontal = TRUE)
```

**Answer**: The height distribution seems roughly symmetric with one outlier!

- Calculate some **numerical summaries** for the **heights**. Briefly summarise your findings. Here you may need to remove `NA`s in the variable before proceed.  

```{r}
### write code here
height_cleaned = math1005$Height[!is.na(math1005$Height)]
summary(height_cleaned)
```


## Write a function in R to compute the population standard deviation

In the lectures, we learned how to write a function in R. Here is your turn, we want to write a function to compute the **population standard deviation** of a data set. Let's name the function `sdpop`. 

- Complete the function in the following.

```{r}
# complete the following function
sdpop <- function(X){
  ### write your code here
  sdpop = sqrt(sum((X-mean(X))^2)/length(X))
  return(sdpop)

}
```

- Calculate the population SD of height and check if it is correct. 

```{r}
### Apply your function sdpop here to height and check its correctness
sdpop(height_cleaned)
sd(height_cleaned) * sqrt((length(height_cleaned)-1)/length(height_cleaned))
```

## Age

- Create a **graphical summary** for the **age** distribution of the cohort. Comment on the distribution, are there any **outliers**? Are there any issues with the data? 

```{r}
### write code here
boxplot(math1005$Age,horizontal = TRUE)
```


- Calculate some **numerical summaries** for the **ages**. Briefly summarise your findings.

```{r}
### write code here
summary(math1005$Age)
```


## Shoe Size

- Create a **graphical summary** for the **shoe size** distribution of the cohort. Comment on the distribution, are there any **outliers**? Are there any issues with the data? **Hint:** Some might response using a different shoe size convention, perhaps using EU instead of US.

```{r}
### write code here
boxplot(math1005$ShoeSize)

convert_to_us <- function(size) {
  # Use ifelse to handle vectors
  us_size <- ifelse(size > 45, size - 33, size)
  return(us_size)
}


boxplot(convert_to_us(math1005$ShoeSize))
```

- Perform any appropriate **data cleaning**. You may also need to remove `NA`s here. 

```{r}
### write code here
shoesize_cleaned = math1005$ShoeSize[!is.na(math1005$ShoeSize)]
shoesize_cleaned
boxplot(convert_to_us(shoesize_cleaned))
```


- Calculate some **numerical summaries** for the **shoe sizes**. Briefly summarise your findings.
  
```{r}
### write code here
summary(shoesize_cleaned)
```

Since Mean is much higher thatn median, the distribution might be right-skewed

# Normal curve

We treat the collected heights as a sample of heights of students in U Syd. We want to fit a normal curve to the histogram of this height sample using the sample mean and sample SD (see Task 2.1). 

## Plot normal curve

The code `curve(dnorm(x, m, s), xlim = c(a, b), add=TRUE)` plots a normal curve with mean m and SD s in the domain $(a,b)$ on the x axis. The option `add=TRUE` let you add a normal curve to an existing plot. See the following for an example. 

```{r}
### here is an example
curve(dnorm(x, 0, 0.5), xlim = c(-10, 5), col = "red")
curve(dnorm(x, -2, 2), xlim = c(-10, 5), col = "blue", add=TRUE)
```

- Now, plot the histogram of height and then plot the normal curve (defined by the sample mean and sample SD) on top of it. Is the normal curve a reasonable approximation to the histogram in this example? 

```{r}
### Write your code here

hist(height_cleaned, probability = TRUE, main = "Histogram with Normal Curve", xlab = "Height")
curve(dnorm(x, mean = mean(height_cleaned), sd = sd(height_cleaned)), 
      col = "blue", 
      xlim = c(140,200), 
      add = TRUE)

```



## Proportion and quantiles under normal curve

- Under the normal curve, are there more students under 160 cm than students taller than 190 cm? Write R code in the following to calculate this. 

```{r}
### Write your code here
#
pnorm(160,mean(height_cleaned),sd(height_cleaned))
1-pnorm(190,mean(height_cleaned),sd(height_cleaned))
```

- Under the normal curve, what is the 23% percentile of the student heights? Write R code in the following to calculate this. 

```{r}
### Write your code here
#
qnorm(0.23,mean(height_cleaned),sd(height_cleaned))
```



# Qualitative Data

## International

- Create a **graphical summary** for the distribution of **domestic and international students** in the cohort. 
  - *Hint: what type of graphical summary would be the most appropriate? You can refer to code in previous lectures and labs*.

```{r}
### write code here

barplot(table(math1005$International))

```

- What is the **proportion** of **international students**? Briefly summarise your findings.
  - *Hint: `math1005$International == "International"` creates a list of `TRUE` and `FALSE` of whether each student is international or not. R can interpret `TRUE` as 1 and `FALSE` as 0.*
```{r}
### write code here

sum(math1005$International == "International")/length(math1005$International)

```

## Gender

- Create a **graphical summary** for the **gender** distribution of the cohort.

```{r}
### write code here
barplot(table(math1005$Gender))
```


- What is the **proportion** of **females**? Briefly summarise your findings

```{r}
### write code here

sum(math1005$Gender=="Female")/length(math1005$Gender)
```


## Country

This task contains some additional visualisation techniques not covered in lectures. These are not mandatory for learning other topics of this unit but can be useful in real life.  

- Create a **barplot** for the **Country** that the students in the cohort were born in. Is it meaningful? Are there any issues with the data? How could this be improved?

```{r}
### write code here
barplot(table(math1005$Country),las=2,cex.names = 0.6)
```

- Make a barplot of the **top 10 countries** among the cohort. Briefly summarise your findings. 
  - *Hint: the `sort` function might be useful!*

```{r}
### write code here

math1005$Country = tolower(math1005$Country)

country_counts = table(math1005$Country)
sorted_country_counts = sort(country_counts,decreasing = TRUE)

top_10_countries = head(sorted_country_counts, 10)

barplot(top_10_countries,cex.names=0.5)
```


- **Extension**: make a **word cloud** of all the majors. Briefly summarise your findings.

You will need to first **install the** `wordcloud` **and** `tm` **packages**. To do this, you can run the first line without the `#`. Once installed, they will remain on your computer and you do not need to run this line ever again. But in each document you want to make a word cloud, you will need to include `library(wordcloud)` to **load the package**. 

```{r, warning = F, message = F}
library(wordcloud)
wordcloud(math1005$Country, min.freq = 1)
```





