---
title: "MATH1062 Lab2 Worksheet"
subtitle: "Introduction to R"
author: "© University of Sydney MATH1062"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: styles.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

# <i class="fa fa-crosshairs"></i> Aim of Lab2 {-}

**Learning Outcome**  
LO3. identify appropriate methods to describe, summarise and visualise a given data set

LO5. apply statistical software such as R to analyse example sets of data

At the end of this Lab, you should be able to:

- Choose what graphical summaries are appropriate for given variables, and interpret the output.
- Produce graphical summaries with R (histogram, boxplot, barplot), and reclassify variables. 

**R commands** `class`, `hist`, `barplot`, `boxplot`, `is.na`, `na.omit`, `sum`, `factor`, ...

# <i class="fa fa-th-list"></i> Summary {-}

Type of Variable | Referred to in R |
-------- | ------- | 
Qualitative / Categorical | `factor`
Quantitative / Numerical | `num`
| |

Type of Data | Type of Graphical Summary | In R |
------------------------ | -------------------------- | -------------- | 
1 Qualitative Variable | Barplot | `barplot()` |
2 Qualitative Variables |  Double (clustered) Barplot |  `barplot()` |
1 Quantitative Variable | Histogram or Boxplot  | `hist()`, `boxplot()`|
2 Quantitative Variables | Scatterplot | `plot()`|
1 Quantitative & 1 Qualitative Variable | Double (comparative) boxplot | `boxplot()`|
| | |


# <i class="fas fa-car"></i> Australian Road Fatalities {-}

Consider the Australian road fatalities from 1989 (a bigger version of the data used in lectures). The data is sourced from [BITRE](https://bitre.gov.au/statistics/safety/fatal_road_crash_database.aspx).

**How to proceed?**

- Check whether you can knit the document without any error. If not, you may need to make sure your `Lab2Worksheet.Rmd` and data file are in the correct folders. 
- Read through each question.
- Try to understand the given code.
- Write short concise answers in this file. 
- Have a go at the explore questions.

# Import the data.

Note: Recall that you need to store your `Lab2Worksheet.Rmd` in your `MATH1062` folder, then download the data from Canvas, store the data in `MATH1062/data` and then run the code. Alternatively, you can store the data in the same folder as `Lab2Worksheet.Rmd`, and remove the `data/` part of the code. You will need to use this method in future projects, when you import your own data.

```{r}
# Read data from csv into R
road = read.csv("AllFatalities.csv",header=T)
```

- Produce a snapshot of the data. 
```{r}
str(road)
```

# Research Questions

## Were there more fatalities on a certain day of the week? 

- Here we consider **1 qualitative variable**: the road fatalities across the *days of the week*.

- First isolate the variable `Dayweek`. Check how R classifies it. Produce a barplot. You may see that the days are in the wrong order. This is because that R automatically sorts factors alphabetically unless we impose an ordering.

```{r}
dayweek = road$Dayweek # isolate the variable Dayweek
class(dayweek) # check how R classify it

### Write your code below to produce the barplot
barplot(table(dayweek), cex.names = 0.8)

```

- We can re-order the categories for `dayweek` and then re-produce a barplot (see lecture slides). What pattern emerges? Suggest possible reasons for it.

```{r}
### Write your code here
dayweek = factor(dayweek, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

barplot(table(dayweek), cex.names = 0.8)

# The highest number of fatalities occur on Friday and Saturday. 

```


In the barplot, you can use the option `las=2` to display the category labels vertically, try it out below. 

```{r}
### Write your code here
barplot(table(dayweek), las = 2)

```

- Summarise what you have discovered.

<br>

## Were there more fatalities in certain age groups?

- Here we consider **1 quantitative variable**: *age*.

- First isolate the variable `Age`. How does R classify it?

```{r}
age = road$Age
class(age)
```

- Change the classification to a quantitative variable (use `as.numeric()`).
```{r}
### Write your code here
age = as.numeric(age)
class(age)
```

- Produce graphical summaries (try boxplot and histogram). What patterns are revealed?
```{r}
### Write your code here
hist(age)
boxplot(age,horizontal = TRUE)

# Right skewed

```

- Summarise what you have discovered.

<br>

## Does gender affect the number of fatalities across age groups?

- Here we consider **1 quantitative variable divided by 1 qualitative variable**.

- Control for gender (using comparative boxplot) - i.e., consider fatalities by age divided by gender.
```{r}
### Write your code here

boxplot(age~Gender, data = road,col = c("lightblue", "lightpink"), horizontal = TRUE)
# (quntitative~qualitative)

```

- Summarise what you have discovered.

The median age of male fatalities is slightly higher compared to females.
Some outliers for male fatalities...

<br>

## Was there any pattern in how buses were involved in fatalities, on different days of the week?

- Here we consider **2 qualitative variables**: the road fatalities across the *days of the week*, cross-classified by *bus involvement*.

- Is there any pattern?

```{r}
### Write your code here
bus_inv = road$Bus_Involvement
data1 = table(bus_inv,dayweek)

barplot(data1,beside = TRUE,legend = rownames(data1),col = c("lightgreen", "lightblue"),cex.names = 0.8)

```

- Summarise what you have discovered.

Most Fatal Accidents Do Not Involve Buses:

The number of fatal road accidents without bus involvement peaks on Friday and Saturday, with a slight drop on Sunday.
<br>

## Data selection

How many deaths were happened at age under 45 and on Saturday? Here you need to consider how to deal with data points with value `NA`

```{r}
### Write your code here
nrow(road[!is.na(road$Age) & road$Age < 45 & road$Dayweek == "Saturday", ])
```

How many deaths were happened at age under 45 and NOT on Saturday? 


```{r}
##### Write your code here
nrow(road[!is.na(road$Age) & road$Age < 45 & road$Dayweek != "Saturday", ])
```


## Were there more fatalities on a certain months of the year? 

- Here we consider **1 qualitative variable**: the road fatalities across the *months of the year*.

Hint: `month.name` contains the months of year sorted in the correct order!

```{r}
month.name
```

```{r}
##### Write your code here
month = factor(road$Month, levels = c(month.name))
barplot(table(month), las=2)

```

- Summarise what you have discovered.

## Was there any pattern between Age and Crash Type?

- Here we consider **1 quantitative variable divided by 1 qualitative variable**. ie consider fatalities by age divided by crash type.

What are the different Crash Types?

```{r}
table(road$Crash_Type)
```

```{r}
### Write your code here
boxplot(age~road$Crash_Type, horizontal = TRUE, col = c("lightgreen","lightblue","red") )

```

- Summarise what you have discovered.

## Was there any pattern in how heavy rigid trucks were involved in fatalities, on different days of the week?

- Here we consider **2 qualitative variables**: the road fatalities across the *days of the week*, cross-classified by *heavy rigid truck involvement*. Note you may need to deal with missing values!

```{r}
### Write your code here

data2 = table(road$Hvy_Rigid_Truck_Involvement,dayweek)
barplot(data2, las = 2,legend = rownames(data2),col = c( "lightblue", "lightgreen", "orange"))

```

- Summarise what you have discovered.

# Challenge

Visualise some of the other variables in the data, and see if you can find any interesting trends.
