---
title: "MATH1062/1005 Lab11 Worksheet"
subtitle: "Two-sample t-test"
author: "© University of Sydney 2024"
date: 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---

# Two Sample t-tests: Common cold

A study is conducted to test whether a new supplement prevents the common cold. A long term study followed 10 individuals over 20 years. In the first 10 years, all individuals were given a placebo capsule. The number of colds over the 10 years were recorded ($y_i$). In the second 10 years, all individuals were given the supplement. The number of colds over the last 10 years were recorded ($x_i$). The pairs of measurements for each individual are given as follows.

supplement ($x_i$): 2, 8, 2, 18, 4, 8, 10, 11, 15, 11

Placebo ($y_i$): 7, 7, 6, 18, 15, 14, 14, 11, 20, 9   

Let $\mu_X$ be the population mean for the number of colds over 10 years while taking the supplement. Let $\mu_Y$ be the population mean for the number of colds over 10 years with the placebo.

1. We will formally conduct a paired (two sample) t-test. 

- The study is interested in whether the supplement has either a positive or negative effect on the number of colds. Hence, state the null and alternative hypotheses.

- Create a vector of the differences between the number of colds in the supplement and placebo groups for each individual.


```{r}
# write code here
```

- Plot a boxplot of the differences. Evaluate the assumptions for a paired (two sample) t-test.

```{r}
# write code here
```

- Manually compute a 95% confidence interval for $\mu_D = \mu_X - \mu_Y$.

```{r}
# write code here
```

- Using the 95% confidence interval, state the conclusion of the test at the 5% significance level.

2. Verify that you computed your confidence interval correctly by using R's inbuilt `t.test()` function on the sample differences.

```{r}
# write code here
```

3. Suppose the researchers of the study accidentally "scrabbled" how the supplement and placebo measurements were paired! 

- Use `sample()` to scramble the measurements in the placebo group and then use `t.test()` to perform a paired t-test on the scrambled differences. 

```{r}
# write code here
set.seed(1234) #Don't touch this line! Its there to make sure we all get the same results :) 

```

- You should have found that the p-value now is higher than before. Why do you think that is? Hint: *patient effect* from lectures.

4. We cannot perform a Welch test since the two samples are dependent. But let's do one anyway and reflect on the results!

- Perform a Welch test using `t.test()`.

```{r}
# write code here
```

- You should have found that the p-value is higher than the paired t-test. Why do you think that is? Hint: it's essentially the same reason at the end of Question 3.






