---
title: "MATH1005 Lab9 Worksheet"
subtitle: "Sample Surveys"
author: "© University of Sydney 2022"
date: 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: 
      # - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

# Aim of Lab9 {-}


At the end of this Lab, you should be able to:

- Use a box model and the Central Limit Theorem to draw inferences about an unknown proportion.
- Demonstrate theory using simulation.

# Same-Sex Marriage Survey

Between 12 September and 7 November 2017 the Australian Government conducted a national postal survey in order to gauge the Australian community's opinion on changing legislation about same-sex marriage.

In the lead-up to the postal survey, various organisations conducted their own polls to infer how the final survey might turn out. [One such poll](https://www.theguardian.com/australia-news/2017/aug/21/most-support-marriage-equality-and-80-plan-to-vote-in-survey-guardian-essential-poll/) returned the following results:

| In favour of proposed change | Against proposed change | Don't know | 
|:----------------------------:|:-----------------------:|:----------:|
|   1036                       | 581                     |   200      |

What can we infer from this about the final results of the official postal survey?

## Simplifying assumptions

Since the final postal survey was to be voluntary, and we ultimately wish to estimate the proportion of participants who favour the proposed change, it is reasonable to

1. discard the "Don't know" counts;
2. model the remaining responses as being like a random sample taken with replacement from a box containing only 0s and 1s, with proportion of 1s equal to $p$ (for some unknown $0<p<1$).

The parameter $p$ is then the unknown "population" proportion of voters in favour of the change. If $p>0.5$, the proposed change to the law will be made by the Government.

**Question**: Do these assumptions seem reasonable? Comment.

## Sample size and observed sample proportion

- **What is the sample size $n$?**

```{r}
n = 1036 + 581
n
```

- **What is the observed sample proportion $\bar x$ ?**
```{r}
obs = 1036/n
obs
```


## Confidence interval

- **Determine a Wilson's 99% confidence interval for $p$ based on the observations in the previous part. Hint:** use `binom.confint()`.

```{r}
# write code here
require(binom)
binom.confint(x = 1036, n, conf.level = 0.99, methods = "wilson")

```

- **Perform a "sanity check" that the endpoints of the interval are such that the observed value $\bar x$ is "right on the edge" of a corresponding 99% prediction interval.**

```{r}
# write code here
box = c(rep(1,1036),rep(0,581))
p = 1036/n

se = sd(box)/ sqrt(n)
se

ex = mean(box)
left = 0.6406926	
right = 0.6707943

a = qnorm(0.995)

CI$lower + a*sqrt(CI$lower*(1-CI$lower)/n)
CI$upper - a*sqrt(CI$upper*(1-CI$upper)/n)
```

## Hypothesis test

- **The observed proportion is greater than 0.5, but is it significantly greater? Answer this question with an appropriate formal hypothesis test.**

```{r}
# write any code here

```


## Results

 The final outcome of the postal survey (i.e. the true population parameter $p$) was 61.6% support. 

- **In light of the results of your calculations above, what can you say about the Essential poll?**


**Note**: you can see the results of the national survey electorate-by-electorate in the following
[ABC Report](http://www.abc.net.au/news/2017-11-15/same-sex-marriage-results-ssm/9145636). 
