---
title: "MATH1062 Lab 5 Worksheet"
subtitle: "Introduction to R"
author: "© University of Sydney MATH1062"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: 
      - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

# <i class="fa fa-crosshairs"></i> Aim of Lab 5 {-}

**Learning Outcome** 

LO1. apply mathematical logic and statistical thinking to solve problems
    
LO3. identify appropriate methods to describe, summarise and visualise a given data set

LO4. identify and apply appropriate methods of inference for a variety of data types

LO5. apply statistical software such as R to analyse example sets of data


# Linear regression 

## Preparation {-}

We start the lab by replicating the regression line and other linear functions we have seen in the lecture. The following commands read Pearson's data set into R as a data frame `data`. We will use Father's height (`fheight`) as the indepdent variable and Son's height (`sheight`) as the dependent variable. 

```{r, echo=F, messages=F,warning=F}
# You may need to install the package UsingR
# install.packages("UsingR")  
suppressMessages(library(UsingR))
library(UsingR) # Loads another collection of datasets
data(father.son) # This is Pearson's data.
data = father.son
str(data)
```

```{r}
### the following code draw a scatter plot with transparent data points 
plot(data$fheight, data$sheight, xlab="Father height (inches)", ylab="Son height (inches)", col=adjustcolor("black", alpha.f = 0.35))
### the following plot a line with intercept a = 10, and slope b = 0.8
abline(a = 10, b = 0.8, col = "purple", lwd = 2) 
```

## Find the slope and intercept of the SD line

Recall that the SD line connects the point of averages $(\bar{x}, \bar{y})$ to $(\bar{x} + \mbox{SD}_x, \bar{y} + \mbox{SD}_y)$. Work out the intercept and slope of the SD line on pen and paper, and then plot the SD line on top of the scatter plot.

**Hint:** the line must pass those two points, so you can first work out the slope, and then identify the intercept. 

```{r}
### Write your code here 

x_mean = mean
sd_y = sd(data$sheight)
sd_x = sd(data$fheight)
slope = sd_y / sd_x

plot(data$sheight~data$fheight)
intercept = mean(data$sheight)  - slope * mean(data$fheight)
abline(a = intercept,b = slope,lwd = 2, col="green")


```



## The regression line

Draw the following three plots. First, plot the regression line on top of the scatter plot. Second, a separate figure will be created that draws the scatter plot of the residual and the independent variable. Using the visualisation to justify if the residuals are homoscedastic? 

```{r}
### write your code here

plot(data$sheight~data$fheight)
model = lm(sheight~fheight,data = data)
abline(model, col="red", lwd=2)

plot(residuals(model)~data$fheight)
abline(h=0, col="green", lwd=2)
model

```


## Sum of squared residuals

- Now write a function that can compute the sum of squared residuals for any line specified by  intercept and slope. Complete your function using the following function definition.

```{r}
sosr <- function(x, y, alpha, beta){
  # x: the independent variable
  # y: the dependent variable
  # alpha: intercept
  # beta:  slope
  #
  ### write your code here
  
  predict_value = alpha + beta * x
  residuals = predict_value - y
  
  sosr = sum(residuals^2)

  return(sosr)
  
}
```

## Comparing the SD line with the regression line and the baseline prediction

Apply the function you wrote to compute the sum of squared residuals of the SD line, the regression line, and the baseline prediction. 

```{r}
### write your code here

## SD line
sd_sosr = sosr(data$fheight, data$sheight, intercept, slope)

regre_sosr = sosr(data$fheight, data$sheight, coef(model)[1], coef(model)[2])

base_sosr = sosr(data$fheight, data$sheight,mean(data$sheight),0)

sd_sosr
regre_sosr
base_sosr


```

## Coefficient of determination

First, compute the coefficient of determination of the linear regression model using your code written above, and then verify that it agrees with $r^2$ (the squared correlation coefficient). What is the proportion of variation in the response variable that can be explained by the linear regression model? 


```{r}
### write your code here
cor(data$sheight,data$fheight)^2

1 - regre_sosr/base_sosr

```

# ChickWeight data 

## Preparation {-}

Now we want to repeat Q1.2, Q1.3, Q1.4, and Q1.5 on the ChickWeight data set. We want to use the `Time` as the independent variable and `weight` as the dependent variable. 


```{r}
x = ChickWeight$Time
y = ChickWeight$weight
plot(x, y)
```

Now we can work with `x` (the time) and `y` (the weight)

## Find the slope and intercept of the SD line


```{r}
### Write your code here

sd_y = sd(y)
sd_x = sd(x)

slope = sd_y / sd_x
intercept = mean(y) - slope * mean(x)

```

## The regression line

Draw the following three plots. First, plot the regression line on top of the scatter plot. Second, a separate figure will be created that draws the scatter plot of the residual and the independent variable. Using the visualisation to justify if the residuals are homoscedastic?

```{r}
### write your code here
model = lm(y~x)
plot(y~x)
abline(model,col="red",lwd = 2)

residuals = y - (model$coefficients[2] * x + model$coefficients[1])
plot(residuals~x)

abline(h=0,col = "green",lwd = 2)

```



## Comparing the SD line with the regression line and the baseline prediction

Apply the function you wrote to compute the sum of squared residuals of the SD line, the regression line, and the baseline prediction.

```{r}
### write your code here

sd_sosr = sosr(x, y, intercept, slope)

regre_sosr = sosr(x, y, coef(model)[1], coef(model)[2])

base_sosr = sosr(x, y,mean(y),0)

sd_sosr
regre_sosr
base_sosr

```



## Coefficient of determination

First, compute the coefficient of determination of the linear regression model using your code written above, and then verify that it agrees with $r^2$ (the squared correlation coefficient). What is the proportion of variation in the response variable that can be explained by the linear regression model?


```{r}
### write your code here

cor(y,x)^2

1 - regre_sosr / base_sosr
```




# MATH1005 data set

## Preparation {-}

Now we want to repeat Q1.2, Q1.3, Q1.4, and Q1.5 on the MATH1005 data set. We want to use the `ShoeSize` as the independent variable and `Height` as the dependent variable. 

- Read in the MATH1005 data set (used the cleaned version)

```{r}
math1005 = read.csv("math1005_cleaned.csv")
```

- From the last lab, we know there are clear outliers in the shoesize, where some responses are probably using a different shoe size convention, perhaps using EU instead of US. This is quite misleading, so perhaps it is best to remove these values (or convert these to the right format).

```{r}
ShoeSize_clean = math1005$ShoeSize[math1005$ShoeSize < 20]  # only keep shoe sizes < 20
Height_clean = math1005$Height[math1005$ShoeSize < 20]  # only keep heights for students with shoe sizes < 20
#
NA_ind = which(is.na(ShoeSize_clean) | is.na(Height_clean))  # index of any NA entries
x = ShoeSize_clean[-NA_ind]
y = Height_clean[-NA_ind]
plot(x, y)
```

Now we can work with `x` (the cleaned shoesize) and `y` (the cleaned height)


## Find the slope and intercept of the SD line

```{r}
### Write your code here

```



## The regression line

Draw the following three plots. First, plot the regression line on top of the scatter plot. Second, a separate figure will be created that draws the scatter plot of the residual and the independent variable. Using the visualisation to justify if the residuals are homoscedastic?

```{r}
### write your code here

```



## Comparing the SD line with the regression line and the baseline prediction

Apply the function you wrote to compute the sum of squared residuals of the SD line, the regression line, and the baseline prediction.

```{r}
### write your code here

```



## Coefficient of determination

First, compute the coefficient of determination of the linear regression model using your code written above, and then verify that it agrees with $r^2$ (the squared correlation coefficient). What is the proportion of variation in the response variable that can be explained by the linear regression model? 


```{r}
### write your code here

```


# Coefficient of determination

Recall that the coefficient of determination of the regression line is
\[
1 - \frac{\text{SSE}}{\text{SST}}, \quad \text{SSE} = \sum_{i = 1}^n (y_i - a - b x_i)^2, \quad \text{SST} = \sum_{i = 1}^n (y_i - \bar{y})^2
\]
where $a$ and $b$ are the intercept and the slope of the regression line, respectively. We want to show that the coefficient of determination can be indeed be written as squared correlation coefficient ($r^2$). The following steps demonstrate one of many ways (with a statistical interpretation) to show this. You can go through these steps and numerically verify those steps using R. **The overall derivation is not examinable, however, going through these steps will be beneficial for understanding the properties of linear regression and correlation. **

## Properties of residuals

Let $\hat{y}_i = a + b x_i$ denote the predicted value of the regression line, and let $\hat{e}_i = y_i - \hat{y}_i$ denote the residual of the regression line. Recall the sum of squared residual function for any line specified by the intercept $\alpha$ and slope $\beta$: 
\[
f(\alpha, \beta) = \sum_{i =1}^n (y_i - \alpha + \beta x_i)^2
\]
and the conditions for the stationary point $(a,b)$ of $f$ (page 52 of topic 4):
\[
\frac{\partial f}{\partial \alpha}(a,b) = 0 \quad \Leftrightarrow \quad  \sum_{i=1}^n \hat{e}_i = 0,
\]
and
\[
\frac{\partial f}{\partial \beta}(a,b) = 0 \quad \Leftrightarrow \quad  \sum_{i=1}^n \hat{e}_i x_i = 0.
\]
In the first step, we want to verify (by derivation) that the sum of product of the residuals and the differences between regression line and the baseline prediction is zero, i.e.,
\[
\sum_{i=1}^n \hat{e}_i (\hat{y}_i - \bar{y}) = 0.
\]

You can also use the `x` and `y` variables from the previous example to verify this numerically. 

```{r}
### Write your code here

```



## Sum of squared differences between regression and baseline

In the second step, we consider the sum of squared differences between the regression line and the baseline prediction: 
\[
\text{SSR} = \sum_{i=1}^n (\hat{y}_i - \bar{y})^2.
\]
We want to verify  \( \text{SST} = \text{SSR} + \text{SSE} \), so that **SSR represents the variability explained by the regression model**.

**Hint:** for each $y_i$, we can first consider the decomposition:
\[
y_i - \bar{y} = (\hat{y}_i - \bar{y}) + (y_i - \hat{y}_i)
\]
where \( \hat{y}_i \) is the predicted value from the regression, and \( y_i - \hat{y}_i \) is the residual (error) for that observation. Then, you can take the sum of squares of both sides and apply the result of the previous step to verify this.


You can continue using the `x` and `y` variables from the previous example to verify this in R. 


```{r}
### write your code here

```


## Correlation between the prediction of the regression line and the response variable

In step three, we want to verify that 
\[
\text{cor}(\hat{y}, y)^2 = \frac{SSR}{SST} = 1 - \frac{\text{SSE}}{\text{SST}} 
\]
so that the correlation coefficient between the predicted values of the regression line $\hat{y}$ and the response variable $y$ gives the coefficient of determination. 

**Hint:** you need to use the definition of the correlation coefficient, and the fact that $\bar{y}$ is also the mean of $\hat{y}$ (why? using the formula of the intercept). 


You can continue using the `x` and `y` variables from the previous example to verify this in R. 


```{r}
### write your code here

```



## Putting the results together

Applying the invariant property of the correlation coefficient and step three, verify that
\[
\text{cor}(x, y)^2 =  \frac{\text{SSR}}{\text{SST}}.
\]


```{r}
l = lm(Orange$circumference ~ Orange$age)
plot(Orange$age, l$residuals)
```




