---
title: "MATH1062/1005 Lab8 Worksheet"
subtitle: "Chance Variability (CLT)"
author: "© University of Sydney 2024"
date: 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: 
      - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

# Aim of Lab 8 {-}


At the end of this Lab, you should be able to:

- Draw a box model for a given random experiment, and calculate the expected value, SE and Normal probabilities for the sum and mean of the draws.
- Simulate the Central Limit Theorem in R.

**R commands** 
`set.seed`, `pnorm`, `sample`, `replicate`


# Box Model Revision

The **box model** models the sum (or mean) of $n$ random draws from a population. The population is the "box", and the collection of draws forms the "sample".

Note: Unless otherwise stated, we assume draws are made with replacement from the box.

(1) The Box

As the box represents a population, we use `popsd(box)`, where `popsd` is defined via:

```{r}
popsd = function(x) sqrt( mean(x^2) - (mean(x)^2) )
```


|Focus|By hand|In R|Symbol |
|--------------|-----|--|-------|
|Mean of the Box|Calculate the the average| `mean(box)`|$\mu$ |
|SD of the Box|Calculate the population sd|  `popsd(box)`|$\sigma$ |

(2) The Sample

We can focus on either the Sum (or Mean) of the sample.

|Focus|Expected Value (EV)|Standard Error (SE)|
|----|----|----|
|Sum of the Sample| $n\mu$  |$\sqrt{n} \sigma$ |
|Mean of the Sample| $\mu$ |  $\frac{\sigma}{\sqrt{n}}$|

## Winnings after of 25 rolls

Scenario: in a game, you roll a fair die once, and if it lands on "1", you win \$4, and otherwise, you lose \$1. Suppose you play this game 25 times. Let $X_1,\ldots,X_n$ denote your winnings each
time (with $n=25$) and $S=X_1+\cdots+X_n$ your total winnings after $n=25$ repetitions of the game.

For each part below, try to work out the calculations "by hand", and then check in R.

- Draw the box model.

- Calculate the mean and SD of the box.

```{r}
## write code here
s = c(4,-1,-1,-1,-1,-1)

mu = mean(s)
omega = popsd(s)
n = 25
```

- Calculate the $E(S)$ and $SE(S)$, the EV and SE of your total winnings after 25 repetitions of the game.

```{r}
## write code here
ev = n * mu
se = sqrt(n) * omega
ev
se
```

- Using the appropriate normal approximation (by the Central Limit Theorem), estimate the chance that you will break even, that is, not lose money. 
  - *Hint: you will need to use R for this, and it may help to draw a diagram.* 

```{r}
## write code here
z = (0 - ev)/se
1 - pnorm(z)
```

- Estimate the chance that you will win more than \$10.

```{r}
## write code here
pnorm(10,ev,se,lower.tail = FALSE)
```

- Estimate the chance that you will lose more than \$10.
  
```{r}
## write code here
z = (-10 - ev)/se
pnorm(z)
```


- Compare these approximations to "true" probabilities using a simulation:
  - simulate a sum of 25 games 1000 (or more) times
  - determine the proportion of simulations where the sum is
    - positive
	- more than \$10
	- less than -\$10
	
Code to obtain the sums in one line is `sums=replicate(1000, sum(sample(box, 25, replace=T)))`.

```{r}
sums=replicate(1000, sum(sample(s, 25, replace=T)) > 10)
mean(sums)

sums=replicate(1000, sum(sample(s, 25, replace=T)) < -10)
mean(sums)

sums=replicate(1000, sum(sample(s, 25, replace=T)) > 0)
mean(sums)
```


You can also compare the shape of the histogram to the normal curve with mean $E(S)$
and SE $SE(S)$ with something like
```{r, eval=F}
sums=replicate(1000, sum(sample(s, 25, replace=T)))
hist(sums, pr=T)
curve(dnorm(x, m=ev, s=se), add=T, lty=2)
```
What do you notice?


## Winnings after 100 rolls

Now consider the same game, but you play 100 times.

- Estimate the chance that you will break even.

```{r}
## write code here
n = 100
ev2 = n * mu
ev2

se2 = sqrt(n) * omega
se2
```

- Estimate the chance that you will win more than \$10.

```{r}
## write code here
z = (10-ev2)/se2
pnorm(z,lower.tail = FALSE)
```

- Estimate the chance that you will lose more than \$10.
  
```{r}
## write code here
z = (-10-ev2)/se2
pnorm(z)
```

- Use a simulation to assess the quality of these normal approximations, and the shape of the histogram. What do you conclude?

```{r}
sums2=replicate(1000, sum(sample(box, 25, replace=T)) > 10)
mean(sums2)
```

## Statistical Thinking

- How do the probabilities differ between 25 plays and 100 plays? Do these results make sense? **Hint:** instead of considering $S$, the total of your winnings, consider $\bar X=S/n$, the average win per roll.

# Simulate the CLT 

## Simulate simple box (symmetric)

Experiment to find out what minimum size number of draws it takes, for the distribution of the sample sum to start looking like a Normal curve.

Method: Use `replicate` to simulate 1000 samples of 10, 30, 100 and 1000 draws from a box, and compare your results.

- Take draws from 0,1 box
```{r}
set.seed(1)
box=c(0,1)
totals1 = replicate(1000, sum(sample(box, 10, rep = T)))
totals2 = replicate(1000, sum(sample(box, 30, rep = T)))
totals3 = replicate(1000, sum(sample(box, 100, rep = T)))
totals4 = replicate(1000, sum(sample(box, 100000, rep = T)))

par(mfrow = c(2,2))
hist(totals1, main = "10 draws")
hist(totals2, main = "30 draws")
hist(totals3, main = "100 draws")
hist(totals4, main = "1000 draws")
```

How many draws do we need to take before the histogram starts looking like a normal distribution?

## Simulate simple box (asymmetric)

- Take draws from 0,1,1,1,1,1,1 box
```{r}
## write code here
set.seed(1)
box = c(0,1,1,1,1,1,1)
totals1 = replicate(1000, sum(sample(box, 10, rep = T)))
totals2 = replicate(1000, sum(sample(box, 30, rep = T)))
totals3 = replicate(1000, sum(sample(box, 100, rep = T)))
totals4 = replicate(1000, sum(sample(box, 1000, rep = T)))

par(mfrow = c(2,2))
hist(totals1, main = "10 draws")
hist(totals2, main = "30 draws")
hist(totals3, main = "100 draws")
hist(totals4, main = "1000 draws")
```

How many draws do we need to take before the histogram starts looking like a normal distribution?

## Simulate your own box

Experiment to find out what minimum size number of draws it takes, for the distribution of the sample sum to start looking like a Normal curve for your own custom box.

```{r}
# write code here

```


## Further experimenting with the CLT

- Experiment with the [Seeing Theory: CLT](https://students.brown.edu/seeing-theory/probability-distributions/index.html) app.
