---
title: "MATH1062 Lab4 Worksheet"
subtitle: "Introduction to R"
author: "© University of Sydney MATH1062"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: 
      - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

# <i class="fa fa-crosshairs"></i> Aim of Lab4 {-}

**Learning Outcome** 

LO1. apply mathematical logic and statistical thinking to solve problems
    
LO3. identify appropriate methods to describe, summarise and visualise a given data set

LO5. apply statistical software such as R to analyse example sets of data

At the end of this Lab, you should be able to:

- Model a variable by a Normal curve, and explain how to find areas under the curve and percentiles.
- Use R to calculate areas under the Normal curve and percentiles.
- Understand correlation coefficient and its properties

**R commands** 
`pnorm`, `qnorm`, `cor`

# <i class="fa fa-th-list"></i> Summary {-}

- It is vital that you draw a *quick sketch* of the Normal curve that you are considering, *before* you attempt any calculations.
- We use R to find an area under the Normal curve. Historically, 'Stats Tables' were used.
- The *area under the curve* up to point $x$, represents the *chance* (or "probability") of getting a value less than or equal to $x$.


# Import data and more plotting tricks

## Data

We continue using the data set collected in MATH1005 in S2, 2022. Make sure you have the data-file `math1005_cleaned.csv`  in your data folder within your MATH1062 folder.

- **Import** the data into a variable called `math1005`:

```{r}
math1005 = read.csv("math1005_cleaned.csv")
str(math1005)
```

## Plotting area under a normal curve

The following code sketches a Normal Curve with mean 40 and SD 15. The red shaded area represents $P(X < 18)$. The blue shaded area represents $P(60 < X < 80)$. 

```{r}
curve(dnorm(x,40,15),from=-20,to=100,ylab="Density",main="N(40,225)")
x = seq(-3.5,3.5,length=1000)*15 + 40
y = dnorm(x,40,15)
y18 = dnorm(18, 40, 15)
polygon(c(min(x), x[x<18], 18, 18), c(0, y[x<18], y18 , 0), col="red")
y60 = dnorm(60, 40, 15)
y80 = dnorm(80, 40, 15)
polygon(c(60, 60, x[x>60&x<80], 80, 80), c(0, y60, y[x>60&x<80], y80, 0), col="blue")
```

# <i class="fas fa-football-ball"></i> Australian men's AFL team and heights

In the Australian Football League (AFL) recruiters tend to look for tall male players. We want to use the heights of male students in MATH1005, S2 2002 as a sample to model the Australian male height. 

## Data modelling

Select the heights of male students from the data set. Plot a histogram of the selected heights. Construct a normal curve to approximate the histogram of male heights in `math1005`. Plot the resulting histogram and then plot the normal curve on top of it. If you have issues with obtaining the mean and SD, you can use mean = $178.84$ and SD = $6.53$,

```{r}
### Write your code here

height_male = math1005[math1005$Gender == 'Male' & !is.na(math1005$Height), "Height"]

# Plot the histogram with density (freq=FALSE means density)
hist(height_male, freq = FALSE, main = "Height Distribution of Males", xlab = "Height",ylim=c(0,0.07))

# Add the normal curve on top of the histogram
curve(dnorm(x, mean(height_male), sd(height_male)), 
      add = TRUE, lwd = 2, col = 'red')

```

## Using R

For each of the following questions, try to use only `pnorm` and `qnorm` to calculate the answer. You can also apply the plotting tricks above to draw the corresponging area under the normal curve. 

1. According to this [article](https://thatsamkid.com/2020/10/26/the-long-and-short-of-it/), the average height of AFL players is 188cm. What is the chance of finding an Australian man taller than 188cm?  

```{r}
### Write your code here
pnorm(188,mean(height_male),sd(height_male),lower.tail = FALSE)
```

2. The tallest AFL player in history is [Aaron Sandilands](https://en.wikipedia.org/wiki/Aaron_Sandilands) at 211cm tall. What is the chance of finding a man of height greater than Aaron Sandilands? 

```{r}
### Write your code here
pnorm(211,mean(height_male),sd(height_male),lower.tail = FALSE)
```

3. What percentage of Australian men are between 170 and 180cm? 

```{r}
### Write your code here
pnorm(180,mean(height_male),sd(height_male))-pnorm(170,mean(height_male),sd(height_male))
```

4. What percentage of Australian men are you taller than?

```{r}
### Write your code here
pnorm(171,mean(height_male),sd(height_male))
```

5. If 90\% of Australian men are below a certain height, what is that height?

```{r}
### Write your code here
qnorm(0.9,mean(height_male),sd(height_male))
```

6. If 40\% of Australian men are above a certain height, what is that height?

```{r}
### Write your code here
qnorm(1-0.4,mean(height_male),sd(height_male))
```

7. What is the interquartile range of heights of Australian men?

```{r}
### Write your code here
summary(height_male)

183 - 175
```

## Standard units

For the above questions 1 to 4, answer the questions again by converting the values to standard units and using the standard normal curve.

1. What is the chance of finding an Australian man taller than 188cm?  

```{r}
### Write your code here
std_unit_1 = (188-mean(height_male))/sd(height_male)
pnorm(std_unit_1,lower.tail = FALSE)
```

2. What is the chance of finding a man of height greater than 211cm? 

```{r}
### Write your code here
std_unit_2 = (211-mean(height_male))/sd(height_male)
pnorm(std_unit_2,lower.tail = FALSE)
```

3. What percentage of Australian men are between 170 and 180cm? 

```{r}
### Write your code here
std_unit_3 = (170-mean(height_male))/sd(height_male)
std_unit_4 = (180-mean(height_male))/sd(height_male)

pnorm(std_unit_4) - pnorm(std_unit_3)

```

4. What percentage of Australian men are you taller than?

```{r}
### Write your code here
std_unit_5 = (171-mean(height_male))/sd(height_male)

```

## Calculating by hand

**Use the 68%-95%-99.7% rule to calculate the following by hand.**

```{r}
sd(height_male)
mean(height_male)
```

- What percentage of Australian men are shorter than 198.43cm? 

178.8377 + 3 * 6.533423 =  198.43 99.7%

- What percentage of Australian men are taller than 165.78cm? 

1 - 0.95

- What percentage of Australian men are between 172.31cm and 191.9cm? 

[mean-sd,mean+sd]
68% - (1-68%) = 36%

- What is the 97.5th percentile of heights of Australian men? 

mean + 2*sd

- What is the 2.5th percentile of heights of Australian men? 

mean - 2*sd

- Write down an interval which contains 95% of the heights. 

[0,mean+2*sd]

[mean-2*sd,1]



# Correlation coefficient

We want to explore the association between shoesize and height. Now we want to use all data points in `math1005`. 

## First attempt

- What is the correlation coefficient of the two variables? - use `cor`. Hint: the argument use = "complete" will ignore NA values.

```{r}
### Write your code here

```

- Produce a **scatterplot** using the ``plot`` function.

```{r}
### Write your code here
plot(math1005$Height,math1005$Shoesize)
```  

- How would you describe the association between shoesize and height? Hint: Some might response using a different shoe size convention, perhaps using EU instead of US. The majority of students reported US shoesize.

**Answer:** 


## Removing outliers

Since the majority of students reported US shoesize, let's discard data points with shoesize > 20 as outliers, and then repeat the above procedure, what is your finding?

```{r}
### Write your code here

```  

**Answer:** 


## More about correlation coefficient

Using the data without outliers, verifying the following properties of the correlation coefficient using R. 

- Verify that the correlation coefficient is not affected by interchanging the variables (**symmetry**).

```{r}
### Write your code here

```

- The conversion from US shoesize to EU shoesize approximately follows the formula `EU ShoeSize` = `US ShoeSize` x 1.27 + 30. Now transform the cleaned shoesizes (assuming they are US sizes) into EU sizes. Then verify that the correlation coefficient is shift and scale invariant.  

```{r}
### Write your code here

```