---
title: "MATH1062/MATH1005 Lab7 Worksheet"
subtitle: "Introduction to R"
author: "© University of Sydney MATH1062/MATH1005"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---

# Aim of Lab 7


At the end of this Lab, you should be able to:

- Calculate expected value and standard error from a box model.
- Understand the significance of expected value and standard error.

# Summary

The **box model** models the sum (or mean) of $n$ random draws from a population. The population is the "box", and the collection of draws forms the "sample".

(1) The Box

As the box represents a population, we use `popsd(box)`, where we write our own function `popsd()` to compute the "population SD" of the box:
```{r}
popsd = function(x) sqrt( mean(x^2)-(mean(x)^2))
```

|Focus|By hand|In R|Symbol |
|--------------|-----|--|-------|
|Mean of the Box|Calculate the the average| `mean(box)`|$\mu$ |
|SD of the Box|Calculate the population sd|  `popsd(box)`|$\sigma$ |

(2) The Sample

We can focus on either the sum (or mean) of the sample.

|Focus|Expected Value |Standard Error |
|----|----|----|
|Sum of the Sample| $n \mu$  |$\sqrt{n} \sigma$ |
|Mean of the Sample| $\mu$ |  $\sigma/\sqrt{n}$|


For each calculation, work out the answer "by hand", and then check your answers in R.

# Simple Binary Box Model 

Model the following scenario: the multiple choice component of your final exam contains 20 questions, each with one correct answer and three incorrect answers. A correct answer is worth 1 mark and an incorrect answer is worth 0 marks. 

Would it be unusual to pass (score at least 50% in) the multiple choice section by randomly guessing? 

1. **Draw the box model**
```{r}
box = c(1,0,0,0)
```

2. **Calculate the mean and SD of the box**

```{r, message = F, warning = F}
# write code here
box_mean = mean(box)
box_popsd = popsd(box)
box_popsd
```

3. **Calculate the $E(S)$ and  $SE(S)$ where $S$ is the sum of draws from the box**

```{r}
# write code here
n = 20

# E(S)
E_S = n * mean(box)
E_S

SE_S = sqrt(n) * popsd(box)
SE_S
```

4. **How many standard errors is the observed value away from the expected value? Would it be unusual to pass by randomly guessing? Hint: assume the box of all possible sums has a normal shape.**

```{r}
# write code here
z = (10 - E_S) / SE_S
z
pnorm(z,lower.tail = FALSE)

pnorm(10,E_S,SE_S,lower.tail = FALSE)
```

# Statistical Thinking 1

If we decided to model the above scenario using the **mean** $\bar X$ instead of the **sum** $S$, which parts would change, and which parts would stay the same? Why?

- **Mean and SD of box**
```{r}
# write code here
box_2 = c(1,0,0,0)

mean(box_2)
popsd(box_2)

```

- **$E(\bar X)$ and $SE(\bar X)$**
```{r}
# write code here

E_x = mean(box)
SE_x = popsd(box)/sqrt(n)

E_x
SE_x
```

- **Number of SE away from expected value**
```{r}
# write code here
z_mean = (0.5 - E_x)/SE_x
z_mean
```

# Statistical Thinking 2

- Would it be more or less unusual to pass the multiple choice section if there were 100 questions instead of 20?

```{r}
# write code here
n_3 = 100
SE_100 = popsd(box)/sqrt(100)

SE_100

z_mean = (0.5 - E_x)/SE_100

z_mean

pnorm(z_mean,lower.tail = FALSE)

```




